---
title: "Wrangling Last.fm data with R"
author: "Daniel Sánchez"
format: html
knitr:
  opts_chunk: 
    echo: true
---

# Intro

This is the main project document, which includes all of the code needed to clean, analyze and visualize my Last.fm data in R.

## Preliminaries

I load the libraries I will need in here. Mostly it is just the main *lastfmR* package which is currently uploaded to [GitHub](https://github.com/ppatrzyk/lastfmR) now, the typical *tidyverse* stuff and *lubridate* for date management.

```{r}
#| label: setup
#| message: false

library(lastfmR) # If not downloaded, use remotes::install_github("ppatrzyk/lastfmR")
library(tidyverse)
library(lubridate)
```

# Getting the data

I will use the main function from the *lastfmR* package, which is `get_scrobbles()`, to get all of the scrobbles (songs listened/streamed) from my Last.fm account.

```{r}
#| label: get-data
scrobbles <- get_scrobbles('damage_inc7', timezone = 'America/Bogota')
scrobbles
```

The `timezone` argument cannot be left as is since it will give me times which are not correct given the timezones. I listened to the majority of my music in Quito, Ecuador, so using that timezone fixes most of the problem, yet I will need to deal with some timezone changes later.

# Fixing timezones

I need to select the data beginning mid-August 2022 (when I travelled to Canada) and have it transformed to Canadian Pacific time. While I could filter out all pre-trip data from Last.fm and bind it to another request to Last.fm but with Vancouver time, I choose to use *lubridate* for a smoother experience.

```{r}
#| label: timezone
scrobbles_can <-
  scrobbles %>% 
  filter(date >= as.POSIXct('2022-09-23 00:00:00', timezone = 'America/Bogota'))

```

We now transform to another timezone:

```{r}
#| label: timezone1
scrobbles_can <-
  scrobbles_can %>% 
  mutate(date = with_tz(scrobbles_can$date, tz = 'Canada/Pacific'))
```

We bind the new data back to the main database, but not without eliminating the wrong timezone data so it works well:

```{r}
#| label: timezone2
scrobbles <-
  scrobbles %>% 
  filter(date < as.POSIXct('2022-09-23 00:00:00', timezone = 'America/Bogota')) %>% 
  bind_rows(scrobbles_can) %>% 
  filter(!is.na(date))
```

# Exporting the data

The data gets exported to a csv to use in other software. 

```{r}
# | label: data-export

write.csv(scrobbles, '../data/scrobbles.csv')

```



